---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidydelta

<!-- badges: start -->
<!-- badges: end -->

Delta Method implementation to estimate standard errors in a {tidyverse} workflow. 

## Installation

You can install the development version of tidydelta from [GitHub](https://github.com/JavierMtzRdz/tidydelta) with:

``` r
remotes::install_github("JavierMtzRdz/tidydelta")
# Or
devtools::install_github("JavierMtzRdz/tidydelta")	
```

## Example

Using tidydelta_m(), the following commands are equivalent:

```{r example}
# Load packages
library(tidydelta)
library(tidyverse)

# Simulate samples
set.seed(547)
x = rnorm(10000, mean = 5, sd = 2)
y = rnorm(10000, mean = 15, sd = 3)

bd <- tibble(x, y)

# Equivalent uses of tidydelta_m()
tidydelta_m(~ y/x,
            conf_lev = .95)

tidydelta_m(~ bd$y/bd$x,
            conf_lev = .95)
bd %>%
  summarise(tidydelta_m(~ y/x,
                        conf_lev = .95))
```

Now, the data frame is divided into samples to compare the transformation of the sample with the estimation of `tidydelta()`. In the real world, you would not need to compute the Delta Method if you have many samples, but it shows how it can be incorporated in a workflow with tidyverse. 

```{r message=FALSE, warning=FALSE, fig.height=3.5, dpi=300}
(result <- bd %>%
  summarise(tidydelta_m(~ x/y,
                        conf_lev = .95)))

ggplot() +
  geom_histogram(data = bd %>%
                   mutate(t = x/y),
                 aes(x = t)) +
  geom_vline(aes(xintercept = result$T_n,
                 color = "T_n")) +
  geom_vline(aes(xintercept = c(result$lower_ci,
                                result$upper_ci),
                 color = "CI"),
             linetype = "dashed") +
  labs(color = element_blank()) +
  theme_minimal() +
  theme(text = element_text(family = "Times New Roman"))
```
